The Evolution of Gene Expression in the Term Placenta of Viviparous Mammals

### 2016-04-29 poster/placenta_viviparous_poster.Rnw
update poster plot margins

\block{Core Placenta Transcriptome}{
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
cptl <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species),]
cptl <- cptl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]
print(ggplot(cptl,
             aes(y=human_name,x=Species))
      + geom_raster(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(),
              axis.text.y = element_text(face="italic"),
              axis.text.x = element_text(angle = 90, hjust = 1, vjust=0,face="italic"))
      )
@ 
\captionof{figure}{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  1:1 human ortholog genes expressed (\ac{FPKM}$ \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied with housekeeping genes
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited} removed
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_transcriptome}
}

@@

common.plot.options <- 
    list(geom_raster(aes(fill = fpkm)),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         ylab("Human Ortholog Name"),
         xlab("Species"),
         theme_grey(),
         theme(legend.position = "top",
               axis.ticks = element_blank(),
               panel.background = element_blank(),
               axis.text.y = element_text(face="italic"),
               axis.text.x = element_text(angle = 90, hjust = 1, vjust=0,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long, aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\captionof{figure}{Expression in $\mathrm{log}_{10}(\mathrm{FPKM})$ of 1:1 \human
  orthologs of Hormone Genes, Galectins, genes implicated in
  pre-eclampsia (PE), and \textit{IGF}-related genes. Empty locations
  indicate no detectable expression and/or no 1:1 \human\ ortholog in
  that species.}
}

>>>>

\block{Core Placenta Transcriptome}{
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
cptl <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species),]
cptl <- cptl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]
print(ggplot(cptl,
             aes(y=human_name,x=Species))
      + geom_raster(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(),
              plot.margin = unit(c(0,0,0,0),"mm"),
              axis.text.y = element_text(face="italic"),
              axis.text.x = element_text(angle = 90, hjust = 1, vjust=0,face="italic"))
      )
@ 
\captionof{figure}{\Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  1:1 human ortholog genes expressed (\ac{FPKM}$ \ge 10$) in
  placentas of all \Sexpr{star.logs[,length(unique(species))]} species
  studied ordered by median expression without housekeeping genes
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}.}
\label{fig:core_placenta_transcriptome}
}

@@

common.plot.options <- 
    list(geom_raster(aes(fill = fpkm)),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         ylab("Human Ortholog Name"),
         xlab("Species"),
         theme_grey(),
         theme(legend.position = "top",
               axis.ticks = element_blank(),
               plot.margin = unit(c(0,0,0,0),"mm"),
               panel.background = element_blank(),
               axis.text.y = element_text(face="italic"),
               axis.text.x = element_text(angle = 90, hjust = 1, vjust=0,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long, aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\captionof{figure}{Expression in $\mathrm{log}_{10}(\mathrm{FPKM})$ of 1:1 \human
  orthologs of Hormone Genes, Galectins, genes implicated in
  pre-eclampsia (PE), and \textit{IGF}-related genes. Empty locations
  indicate no detectable expression and/or no 1:1 \human\ ortholog in
  that species.}
}
